---
title: "Analysis"
author: "Elizabeth Thompson"
date: "11/18/2023"
output:
  html_document: default
  pdf_document: default
---
```{r}
library(forecast)
library(tseries)
library(fUnitRoots)
library(astsa)
library(readr)
library(readxl)
library(TSA)
library(MASS)
```

```{r}
data <- read.csv("energy_final.csv", header = TRUE)
tempdata <- read_xlsx("Temperature Data.xlsx")
```


```{r}
data <- as.vector(data$energy_used)
head(data)
write.table(data, file = "energy.txt")

tempdata <- as.vector(tempdata$Avg)
head(tempdata)
write.table(tempdata, file = "avgtemp.txt")
```

```{r}
plot.ts(data)
acf(data, lag.max = 20)
pacf(data, lag.max = 20)
```
Take a difference? Try ADF for lag one, since we are considering an AR(2) based on PACF
Difference is rejected, so no need to take a difference
```{r}
adfTest(data,1 ,type=c("c"))
adfTest(data,1 ,type=c("ct")) 
```


```{r}
mod1= arima(data, order = c(7, 0, 2))
acf(residuals(mod1))
pacf(residuals(mod1))
print(AIC(mod1))
print(BIC(mod1))

residuals_mod1 <- residuals(mod1) 
lags <- c(6, 12, 18, 24)
test_results <- lapply(lags, function(lag) {
  Box.test(residuals_mod1, lag=lag, type="Ljung-Box")
})

p_values <- sapply(test_results, function(x) x$p.value)
diag_table <- data.frame(Lag = lags, P_Value = p_values)

print(diag_table, digits = 16)
```


```{r}
mod2= arima(data, order = c(8, 0, 0))
acf(residuals(mod2))
pacf(residuals(mod2))
print(AIC(mod2))
print(BIC(mod2))

residuals_mod2 <- residuals(mod2) 
lags <- c(6, 12, 18, 24)
test_results <- lapply(lags, function(lag) {
  Box.test(residuals_mod2, lag=lag, type="Ljung-Box")
})

p_values <- sapply(test_results, function(x) x$p.value)
diag_table <- data.frame(Lag = lags, P_Value = p_values)

print(diag_table, digits = 16)
```

```{r}
sqrtdata = sqrt(data)
sqrttemp = sqrt(tempdata)
```

```{r}
adfTest(sqrtdata,2 ,type=c("c"))
adfTest(sqrtdata,2 ,type=c("ct")) 
```
Try Transfer Function Model 
Yt = Energy data
Xt = Temperature data

```{r}
sqrtdata_ts = ts(sqrtdata)
sqrttemp_ts = ts(sqrttemp)
```

How do the datasets look next to each other?
```{r}
plot.ts(data, col = "blue")

# Add the second time series
plot.ts(tempdata, col = "red")
par(mfrow = c(1,2))
```


```{r}
plot.ts(sqrtdata_ts, sqrttemp_ts)
par(mfrow = c(1,2))
pacf(sqrtdata_ts, sqrttemp_ts)
acf(sqrtdata_ts, sqrttemp_ts)
```


```{r}
tfmod = arimax(sqrttemp_ts, order=c(2,0,2), xreg=sqrtdata_ts, xtransf=sqrtdata_ts, transfer=list(c(1,0)))

# Check model diagnostics
checkresiduals(tfmod)

# Summary of the model
summary(tfmod)

residuals_tfmod <- residuals(tfmod) 
lags <- c(6, 12, 18, 24)
test_results <- lapply(lags, function(lag) {
  Box.test(residuals_tfmod, lag=lag, type="Ljung-Box")
})

p_values <- sapply(test_results, function(x) x$p.value)
diag_table <- data.frame(Lag = lags, P_Value = p_values)

print(diag_table, digits = 16)
```

There is feedback in our Transfer Function Model
```{r}
checkresiduals(tfmod)

tempdata_ts = ts(tempdata)
sqrtdata_ts = ts(sqrtdata)

# Compute and plot cross-correlation
ccf_result = ccf(sqrtdata_ts, tempdata_ts, lag.max=10, plot = FALSE)

plot(ccf_result, main="Cross-Correlation between x and y", ylab="CCF")
```

```{r}
final_mod = arima(sqrtdata, order = c(1, 0, 1))
residuals_model <- residuals(final_mod)

# Generate the Q-Q plot
qqnorm(residuals_model)
qqline(residuals_model, col = "red")

# Optionally, you can add a title
title("Normal Q-Q Plot of Residuals")
```
```{r}
detectAO(final_mod)
detectIO(final_mod)
final_mod2=arimax(sqrtdata,order=c(1,0,1),io=c(110,125))
detectAO(final_mod2)
detectIO(final_mod2)
```

m1.D=arima(seriesD,order=c(1,0,0))
  > m1.D
  > detectAO(m1.D); detectIO(m1.D)
  > m2.D=arimax(seriesD,order=c(1,0,0),io=c(217))
  > m2.D

